Study of Higgs plus single top production using events with a same sign dimuon at the Large Hadron Collider

### 2019-12-11 presentacion_hiram/presentacion-tesis.tex
n

 \author{\scriptsize
 	Universidad de Sonora\\
 	\vspace{1px}
 	Divisi\'on de Ciencias Exactas y Naturales\\
 	\vspace{1px}
 	Departamento de Investigaci\'on en F\'isica\\
 	   \vspace{1px}
   \textbf{L.F. Hiram Ernesto Dami\'an}\\
   Defensa de tesis para obtener el grado de \\
      Maestr\'ia en Ciencias (F\'isica)\\
         \vspace{1px}
   Director:\\
   \vspace{1px}
   Jos\'e Feliciano Ben\'itez Rubio\\
\vspace{1px}
11 de Diciembre de 2019}
 \title{Study of Higgs plus single top production using events with a same sign dimuon  at the Large Hadron Collider}
 %\subtitle{}
 \date{}
 \logo{\includegraphics[scale=0.1]{unison-logo.png}}
 \newcommand{\nologo}{\setbeamertemplate{logo}{}}
 %\subject{}
 \newcommand{\subf}[2]{%
   {\small\begin{tabular}[t]{@{}c@{}}
     #1\\#2
   \end{tabular}}%
 }
 %\setbeamercovered{transparent}
 %\setbeamertemplate{navigation symbols}{}

@@

\begin{frame}
\frametitle{Motivation  for single top Higgs (tH)}
\begin{itemize}
\item Coupling measurement is essential to establish the nature of the Higgs
\item The exploration of Higgs production on the $tH$ channel is subject relatively new. Measurements of CMS and ATLAS are compatible with SM predictions.
\item The $tH$ study explores the relative sign of top-Higgs  and W-Higgs couplings.\\ Small deviations from SM predictions could be associated with physics beyond the
standard model (BSM) such as Supersymmetry.
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Electroweak SM Lagrangian}
	\begin{figure}[ht!]
		\centering
		\begin{tikzpicture}[scale=1.2]
		\begin{axis}[
		hide axis,
		%axis lines=middle,
		%      axis on top,
		%      axis line style={blue,dashed,thick},
		%      ymin=-2,ymax=2,
		%      xmin=-2,xmax=2,
		%      zmin=-2,zmax=2,
		samples=30,
		domain=0:360,
		y domain=0:1.25,clip=false
		]
		\addplot3 [surf, shader=flat, draw=black, fill=white, z buffer=sort]
		({sin(x)*y}, {cos(x)*y}, {(y^2-1)^2});
		\draw[blue,thick,dashed] (axis cs:0,0,0) -- (axis cs:1,0,0)
		node[below,font=\footnotesize]{$\Phi_{\text{IM}}$};
		\draw[blue,thick,-stealth] (axis cs:1,0,0) -- (axis cs:1.3,0,0)
		node[above,font=\footnotesize]{};
		\draw[blue,thick,dashed] (axis cs:0,0,0) -- (axis cs:0,-1,0)
		node[left=2mm,font=\footnotesize]{$\Phi_{\text{RE}}$};
		\draw[blue,thick,-stealth] (axis cs:0,-1,0) -- (axis cs:0,-1.5,0)
		node[right=1mm,font=\footnotesize]{};
		\draw[blue,thick,dashed] (axis cs:0,0,0) -- (axis cs:0,0,1)
		%node[left=2mm,font=\footnotesize]{$\phi_{\text{RE}}$}
		;
		\draw[blue,thick,-stealth] (axis cs:0,0,1) -- (axis cs:0,0,1.3)
		node[right,font=\footnotesize]{$V(\Phi)$};
		\end{axis}
		\end{tikzpicture}
		\caption*{\small Higgs potential $V(\Phi^\dagger \Phi)$ for $\mu^2<0$ and $\lambda>0$, also called the mexican hat potential.}
		\label{hat}
	\end{figure}
\end{frame}
\begin{frame}
\frametitle{SM Lagrangian}
Yukawa lagrangian
\begin{align}
\mathcal{L}_\text{yukawa}=\sum_{m,n}^{3}  \Gamma^u_{mn}\bar{q}_{m\text{,}L} \tilde{\Phi} u_{n\text{,}R}%+\Gamma^d_{mn}\bar{q}_{m\text{,}L} \Phi d_{n\text{,}R}+\Gamma^e_{mn}\bar{l}_{m\text{,}L} \Phi e_{n\text{,}R}+h.c
\end{align}
\scriptsize{where h.c is hermitian conjugate. m and n are flavors of quarks and leptons.
\\
The matrices $\Gamma_{mn}$ describe the so called Yukawa couplings between higgs doublet $\phi$ and the fermions.\\
Choosing}
\begin{align*}
\Phi=-\frac{1}{2}\left(\begin{array}{c}
0 \\
v+h
\end{array} \right) \label{vmin}
\end{align*}
\scriptsize{
By using \ref{vmin} it on $\mathcal{L}_\text{yukawa}$, it is obtained for $u$ quark
\begin{align*}
\frac{\Gamma^u_{uu}v}{\sqrt{2}}(\bar{u_L}u_R+\bar{u_R}u_L)
\end{align*}
from which the masses for the fermions shown to be proportional to the Yukawa couplings
\begin{align*}
m_u=-\frac{\Gamma^u_{uu}v}{\sqrt{2}}
\end{align*}}
\end{frame}

@@

\item L is called luminosity and it is measured in cm$^{-2}$s$^{-1}$.\cite{6}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Higgs production Cross section}
\begin{table}
	\caption*{Higgs boson production cross sections  in pp collisions for $\sqrt{s}=13$TeV  (in
		pico barn).Integrated luminosity of 35.9 fb$^{-1}$ for Run 2\footnotemark}
	\begin{tabular}{|c|c|c|}
		\hline
		Production mechanism &
		$\sigma$ (picobarns pb)
		&Number of events \\
		\hline
		ggF &
		48.93 &
		1756587\\
		\hline
		VBF &
		3.78&
		135702\\
		\hline
		WH & 1.35 & 48465\\
		\hline
		ZH &0.88 & 31592\\
		\hline
		t$\bar{t}$H &
		0.50&
		18183\\
		\hline
		$tH$ &
		0.074&
		2665\\
		\hline
	\end{tabular}
\end{table}
\footnotetext[1]{\tiny{Data taken from The cern collaborarion “Higgs Physics the HL-LHC and HE-LHC” 2019, CERN-LPCC-2018-04}}
\end{frame}

@@

\begin{frame}
\frametitle{Higgs Branching ratios per channel}
\begin{table}
	\caption*{SM Higgs boson branching ratios for  $M_H$ =125 GeV}
\begin{center}
\begin{tabular}{|c|c|}
	\hline
	Higgs decay & Branching ratio (BR)\\
	\hline
	H $\rightarrow$ b$\bar{b}$ &
	$50.82\%$ \\
	\hline
	H $\rightarrow$ $W^+W^-$ &
	$21.5\%$ \\
	\hline
	H $\rightarrow$ $\tau^+ \tau^-$ &
	$6.27\%$\\
\hline
	H $\rightarrow$ ZZ &
$2.61\%$\\
\hline
	H $\rightarrow$ $\gamma\gamma$ &
$0.227\%$\\
\hline
	H $\rightarrow$ Z$\gamma$ &
$0.153\%$\\
\hline
	H $\rightarrow$ $\mu^+\mu^-$ &
$0.0217\%$\\
\hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{frame}
\subsection{CMS dectector}
\frametitle{CMS detector}
\begin{center}
	\begin{figure}
		\includegraphics[scale=0.08]{figures/cms.png}
		\caption*{Compact muon solenoid}
	\end{figure}
\end{center}
\end{frame}

@@

\begin{multicols}{2}
	\small{
\begin{itemize}
\item $t\rightarrow W$$^+$b$\rightarrow$ $l^+ \nu_\mu b$
\item $H\rightarrow$ $W^+$$W^-$

\begin{list}{--}{}
\item  \small{$W\rightarrow$ $l^\pm \nu$, where $l$ can be $\mu^\pm$,$e^-$  $\tau^\pm$} 
\item   \small{$W\rightarrow$ $q\bar{q}$}
%\item \small{$W$ bosons decay leptonically resulting in a signature of two same-sign leptons }
\end{list}
\item A light-flavor quark (Forward jet)
\item  A b-jet  
\end{itemize}
}
\columnbreak
\begin{figure}
	\centering
	\includegraphics[scale=0.5]{figures/jet.png}
\end{figure}
\end{multicols}
\end{frame}

@@

\begin{frame}
\subsection{BDT}
\frametitle{Multivariable discriminator algorithm}
\small{
	A decision tree takes a set of input features and splits input data recursively based on
	those features.
	Boosting is a method of combining many weak learners (trees) into a strong classifier. The features can be a mix of categorical and continuous data.\\
	\vspace{10px}
	BDT Variables}
\tiny{
	\begin{itemize}
		\item	Trailing lepton $p_{t}$
		\item 	Total charge of tight leptons
		\item 	min $\Delta$R (lepton pairs)
		\item 	$\Delta\phi$ between highest $p_t$ lepton pair
		\item 	Number of jets with |$\eta$| < 2.4
		\item	Number of non b-tagged jets with |$\eta$| > 1.0
		\item	Maximum |$\eta$| for jets
		\item	$\Delta\eta$ (most forward light jet, closest lepton)
		\item	$\Delta\eta$ (most forward light jet, hardest loosely b-tagged jet)
		\item	$\Delta\eta$ (most forward light jet, 2nd hardest loosely b-tagged jet)
\end{itemize} }
\end{frame}

@@

{\nologo
\begin{frame}
	\begin{figure}[ht]
		\centering
		\includegraphics[scale=0.5]{figures/bdt1}
		\caption*{BDT training}
	\end{figure}
\end{frame}}

{\nologo
\begin{frame}
\subsection{Previous results}
\frametitle{Previous results for $t\bar{t}H$+$tH$ production}
\begin{minipage}[c]{0.45\textwidth}
\begin{figure}
	\centering
	\includegraphics[scale=0.3]{figures/histo.png}
	\caption*{\tiny{Pre-fit classifier outputs for same sign dimuon, for training against $t\bar{t}V$ }}
\end{figure}
\end{minipage}
\hfill
\begin{minipage}[c]{.45\textwidth}
	\small
\begin{table}
	\centering
	\caption*{\tiny{Event yields for signal and backgrounds after the event selection for a integrated luminosity of 35.9 fb$^{-1}$. The uncertainties of yields include statistical and systematic \protect \footnotemark} }
	\begin{tabular}{cc}
		\hline
		Process & Number of events \\
		\hline
		$t\bar{t}W$ & 68 $\pm$ 10 \\
		$t\bar{t}Z$ & 25.9 $\pm$ 3.9\\
		$WZ$ & 15.1$\pm$7.7\\
		Rares & 20.9 $\pm$ 4.9\\
		Fakes & 80.9 $\pm$9.4\\
		$t\bar{t}H$ & 24.2 $\pm$ 2.1 \\
		\hline
		$tH$ (SM) & 2.14 $\pm$ 0.13\\
		$tH$ ($k_t=-1$) &26.2 $\pm$ 2.2
	\end{tabular}	
	\label{tth-table}
\end{table}

@@

{\nologo
\begin{frame}
\frametitle{Distribution of events in the BDT discriminant}
\begin{minipage}[b]{.48\textwidth}
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{figures/kos.png}
\end{figure}
\end{minipage}
\hfill
\begin{minipage}[b]{.48\textwidth}
	\begin{figure}
		\centering
		\includegraphics[width=\textwidth]{figures/kos2.png}
	\end{figure}
\end{minipage}
\begin{center}
\small{SM signal and backgrounds (Left) and $k_t$=-1 model (Right).}
\end{center}
\end{frame}
}

>>>>

\setbeamercolor*{frametitle}{bg=white}

 \author{\scriptsize
 	Universidad de Sonora\\
 	\vspace{1px}
 	Divisi\'on de Ciencias Exactas y Naturales\\
 	\vspace{1px}
 	Departamento de Investigaci\'on en F\'isica\\
 	   \vspace{5px}
   \textbf{L.F. Hiram Ernesto Dami\'an}\\
   Defensa de tesis para obtener el grado de \\
      Maestr\'ia en Ciencias (F\'isica)\\
         \vspace{5px}
   Director:\\
   \vspace{1px}
   Jos\'e Feliciano Ben\'itez Rubio\\
\vspace{5px}
11 de Diciembre de 2019}
 \title{Study of Higgs plus single top production using events with a same sign dimuon  at the Large Hadron Collider}
 %\subtitle{}
 \date{}
 \logo{\includegraphics[scale=0.1]{unison-logo.png}}
 \newcommand{\nologo}{\setbeamertemplate{logo}{}}
 %\subject{}
 \newcommand{\subf}[2]{%
   {\small\begin{tabular}[t]{@{}c@{}}
     #1\\#2
   \end{tabular}}%
 }
 %\setbeamercovered{transparent}
 %\setbeamertemplate{navigation symbols}{}

@@

\begin{frame}
\frametitle{Motivation  for single top Higgs (tH)}
\begin{itemize}
\item Coupling measurement is essential to establish the nature of the Higgs
\item The exploration of Higgs production on the $tH$ channel is subject relatively new. Measurements of CMS and ATLAS are compatible with SM predictions.
\item The $tH$ study explores the relative sign of top-Higgs  and W-Higgs couplings.\\ Small deviations from SM predictions could be associated with physics beyond the
standard model (BSM)..
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Electroweak SM Lagrangian}
	\begin{figure}[ht!]
		\centering
		\begin{tikzpicture}[scale=1.2]
		\begin{axis}[
		hide axis,
		%axis lines=middle,
		%      axis on top,
		%      axis line style={blue,dashed,thick},
		%      ymin=-2,ymax=2,
		%      xmin=-2,xmax=2,
		%      zmin=-2,zmax=2,
		samples=30,
		domain=0:360,
		y domain=0:1.25,clip=false
		]
		\addplot3 [surf, shader=flat, draw=black, fill=white, z buffer=sort]
		({sin(x)*y}, {cos(x)*y}, {(y^2-1)^2});
		\draw[blue,thick,dashed] (axis cs:0,0,0) -- (axis cs:1,0,0)
		node[below,font=\footnotesize]{$\Phi_{\text{IM}}$};
		\draw[blue,thick,-stealth] (axis cs:1,0,0) -- (axis cs:1.3,0,0)
		node[above,font=\footnotesize]{};
		\draw[blue,thick,dashed] (axis cs:0,0,0) -- (axis cs:0,-1,0)
		node[left=2mm,font=\footnotesize]{$\Phi_{\text{RE}}$};
		\draw[blue,thick,-stealth] (axis cs:0,-1,0) -- (axis cs:0,-1.5,0)
		node[right=1mm,font=\footnotesize]{};
		\draw[blue,thick,dashed] (axis cs:0,0,0) -- (axis cs:0,0,1)
		%node[left=2mm,font=\footnotesize]{$\phi_{\text{RE}}$}
		;
		\draw[blue,thick,-stealth] (axis cs:0,0,1) -- (axis cs:0,0,1.3)
		node[right,font=\footnotesize]{$V(\Phi)$};
		\end{axis}
		\end{tikzpicture}
		\caption*{\small Higgs potential $V(\Phi^\dagger \Phi)$ for $\mu^2<0$ and $\lambda>0$, also called the mexican hat potential.}
		\label{hat}
	\end{figure}
\end{frame}
\begin{frame}
\frametitle{SM Lagrangian}
Yukawa lagrangian
\begin{align}
\mathcal{L}_\text{yukawa}=\sum_{m,n}^{3}  \Gamma^u_{mn}\bar{q}_{m\text{,}L} \tilde{\Phi} u_{n\text{,}R}%+\Gamma^d_{mn}\bar{q}_{m\text{,}L} \Phi d_{n\text{,}R}+\Gamma^e_{mn}\bar{l}_{m\text{,}L} \Phi e_{n\text{,}R}+h.c
\end{align}
\scriptsize{where h.c is hermitian conjugate. m and n are flavors of quarks and leptons.
\\
The matrices $\Gamma_{mn}$ describe the so called Yukawa couplings between Higgs doublet $\phi$ and the fermions.\\
Choosing}
\begin{align*}
\Phi=-\frac{1}{2}\left(\begin{array}{c}
0 \\
v+h
\end{array} \right) \label{vmin}
\end{align*}
\scriptsize{
By using \ref{vmin} it on $\mathcal{L}_\text{yukawa}$, it is obtained for $u$ quark
\begin{align*}
\frac{\Gamma^u_{uu}v}{\sqrt{2}}(\bar{u_L}u_R+\bar{u_R}u_L)
\end{align*}
from which the masses for the fermions shown to be proportional to the Yukawa couplings
\begin{align*}
m_u=-\frac{\Gamma^u_{uu}v}{\sqrt{2}}
\end{align*}}
\end{frame}

@@

\item L is called luminosity and it is measured in cm$^{2}$s$^{-1}$.\cite{6}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Higgs production Cross section}
\begin{table}
	\caption*{Higgs boson production cross sections  in pp collisions for $\sqrt{s}=13$TeV  (in
		pico barn).Integrated luminosity of 35.9 fb$^{-1}$ for Run 2\footnotemark}
	\begin{tabular}{|c|c|c|}
		\hline
		Production mechanism &
		$\sigma$ (picobarns pb)
		&Number of events \\
		\hline
		ggF &
		48.93 &
		1756587\\
		\hline
		VBF &
		3.78&
		135702\\
		\hline
		WH & 1.35 & 48465\\
		\hline
		ZH &0.88 & 31592\\
		\hline
		t$\bar{t}$H &
		0.50&
		18183\\
		\hline
		$tH$ &
		0.074&
		2665\\
		\hline
	\end{tabular}
\end{table}
\footnotetext[1]{\tiny{Data taken from The CERN collaboration “Higgs Physics the HL-LHC and HE-LHC” 2019, CERN-LPCC-2018-04}}
\end{frame}

@@

\begin{frame}
\frametitle{Higgs Branching ratios per channel}
\begin{table}
	\caption*{SM Higgs boson branching ratios for  $M_H$ =125 GeV}
\begin{center}
\begin{tabular}{|c|c|}
	\hline
	Higgs decay & Branching ratio (BR)\\
	\hline
	H $\rightarrow$ b$\bar{b}$ &
	$58.4\%$ \\
	\hline
	H $\rightarrow$ $W^+W^-$ &
	$21.5\%$ \\
	\hline
	H $\rightarrow$ $\tau^+ \tau^-$ &
	$6.27\%$\\
\hline
	H $\rightarrow$ ZZ &
$2.61\%$\\
\hline
	H $\rightarrow$ $\gamma\gamma$ &
$0.227\%$\\
\hline
	H $\rightarrow$ Z$\gamma$ &
$0.153\%$\\
\hline
	H $\rightarrow$ $\mu^+\mu^-$ &
$0.0217\%$\\
\hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{frame}
\subsection{CMS detector}
\frametitle{CMS detector}
\begin{center}
	\begin{figure}
		\includegraphics[scale=0.08]{figures/cms.png}
		\caption*{Compact muon solenoid}
	\end{figure}
\end{center}
\end{frame}

@@

\begin{multicols}{2}
	\small{
\begin{itemize}
\item $t\rightarrow W^+$$^+$b$\rightarrow$ $l^+ \nu_\mu b$
\item $H\rightarrow$ $W^+$$W^-$

\begin{list}{--}{}
\item  \small{$W\rightarrow$ $l^\pm \nu$, where $l$ can be $\mu$,$e^-$  $\tau$} 
\item   \small{$W\rightarrow$ $q\bar{q'}$}
%\item \small{$W$ bosons decay leptonically resulting in a signature of two same-sign leptons }
\end{list}
\item A light-flavor quark (Forward jet)
\item  A b-jet  
\end{itemize}
}
\columnbreak
\begin{figure}
	\centering
	\includegraphics[scale=0.5]{figures/jet.png}
\end{figure}
\end{multicols}
\end{frame}

@@

\begin{frame}
\subsection{BDT}
\frametitle{Multivariable discriminator algorithm}
\small{
	A decision tree takes a set of input features and splits input data recursively based on
	those features.
	Boosting is a method of combining many weak learners (trees) into a strong classifier. The features can be a mix of categorical and continuous data.\\
	\vspace{10px}
	BDT Variables}
\tiny{
	\begin{enumerate}
		\item	$p_t$ of the sub leading leptons
		\item 	jets with largest value of $|\eta|$
		\item 	Number of jets with high $p_t$ and |$\eta$| < 2.4
		\item	Number of jets with |$\eta$| > 1.0
		\item	$\Delta\eta$ (most forward jet, closest lepton)
		\item	$\Delta\eta$ (most forward jet and b-tagged jet)
		\item 	min $\Delta$R (lepton pairs)
		\item 	$\Delta\phi$ between highest $p_t$ lepton pair
	%	\item	$\Delta\eta$ (most forward light jet, 2nd hardest loosely b-tagged jet)
\end{enumerate} }
\end{frame}

@@

{\nologo
\begin{frame}
	\begin{figure}[ht]
		\centering
		\includegraphics[scale=0.48]{figures/bdt-8var}
		\caption*{\scriptsize BDT training}
	\end{figure}
\end{frame}}

{\nologo
\begin{frame}
\subsection{Event distributions and yields}
\frametitle{Event distributions and yields}
\begin{minipage}[c]{0.45\textwidth}
\begin{figure}
	\centering
	\includegraphics[scale=0.18]{figures/kos.png}
	\caption*{\tiny{Pre-fit classifier outputs for same sign dimuon }}
\end{figure}
\end{minipage}
\hfill
\begin{minipage}[c]{.45\textwidth}
	\small
\begin{table}
	\centering
	\caption*{\tiny{Event yields for signal and backgrounds after the event selection for a integrated luminosity of 35.9 fb$^{-1}$. The uncertainties of yields include statistical and systematic \protect \footnotemark} }
	\begin{tabular}{cc}
		\hline
		Process & Number of events \\
		\hline
		$t\bar{t}W$ & 68 $\pm$ 10 \\
		$t\bar{t}Z$ & 25.9 $\pm$ 3.9\\
		$WZ$ & 15.1$\pm$7.7\\
		Rares & 20.9 $\pm$ 4.9\\
		Fakes & 80.9 $\pm$9.4\\
		$t\bar{t}H$ & 24.2 $\pm$ 2.1 \\
		\hline
		$tH$ (SM) & 2.14 $\pm$ 0.13\\
		$tH$ ($k_t=-1$) &26.2 $\pm$ 2.2
	\end{tabular}	
	\label{tth-table}
\end{table}

@@
